---
title: "For Presentation"
author: "Lucy Grigoroff"
date: "2023-05-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
functionDir <- "./R/"
functionList <- list.files(path = functionDir)
for(i in functionList){
  source(paste0(functionDir,i))
}
```

```{r}
ann<-local(get(load("~/OneDrive - Murdoch University/datasets/covid19/bioGune/dataElements/covid19_bioGune_PLA_ANN.daE")))
aaMeta1<-ann@obsDescr[[1]]

aa<-local(get(load(file.path("~/OneDrive - Murdoch University/datasets/covid19/bioGune/dataElements/covid19_bioGune_PLA_SPC.daE"))))
aaData<-data.frame(apply(aa@.Data,2,as.numeric)) #lipoproteins
aaMeta2<-aa@obsDescr[[1]] #metadata from covid19_mauritius_PLA_LIPO.daE file

# match Amino Acids and both Annotations
aaMeta1<-aaMeta1[which(aaMeta1$sampleID %in% aaMeta2$sampleID),]
aaData<-aaData[which(aaMeta2$sampleID %in% aaMeta1$sampleID),]
aaMeta2<-aaMeta2[which(aaMeta2$sampleID %in% aaMeta1$sampleID),]

aaMetaF<-merge(aaMeta1,
                  aaMeta2,
                  by.x ="sampleID",by.y = "sampleID")
rm(aaMeta1, aaMeta2, aa, ann)

BIOspcglyc<-cbind(aaData, aaMetaF)
rm(aaData, aaMetaF)

which(BIOspcglyc[, 1:8]=="Inf")
which(is.na(BIOspcglyc[, 1:8]))
which(is.na(BIOspcglyc$age))
unique(BIOspcglyc$sex)

unique(BIOspcglyc$group)
BIOspcglyc$covid_status<-ifelse(BIOspcglyc$group =="COVID-pos","covid", ifelse(BIOspcglyc$group=="preCOVID","control", "control"))
```

```{r}
loadPackages(myPackages = c("ggplot2",
                            "ggpmisc",
                            "ggplotly",
                            "plotly",
                            "prcomp", 
                            "tidyverse", 
                            "ggpubr",
                            "GGally",
                            "cowplot",
                            "gridExtra",
                            "rlist",
                            "ggrepel"))
```
#generate principal component results
Uses prcomp (uses svd method as opposed to eigenvalue method that princomp uses)
```{r}
#1) create pcresults object
a <- pcResults(BIOspcglyc[,1:8], annotation = BIOspcglyc[,-1:-8])
```

#decide on a threshold
Can change the cutoff but the default is 99%. 
```{r}
#2) create screeCumulativeThresholdObject
b <- screeCumulativeThreshold(pcResultsObject = a, cutoff = 95)
```

#pull out plots
```{r}
b$plots$combinedScreeCumulative
b$plots$screeCumulativeThreshold
b$plots$cumulativeVariance
b$plots$thresholdTable
```

#PCA grid with ellipses
```{r}
#3) gridplot tests various combinations
##working
pcaGrid(screeCumulativeThresholdObject = b, CO = "covid_status", COtitle = "covid", gridTitle = "only colour")

pcaGrid(screeCumulativeThresholdObject = b, CO = "covid_status", SZ = "age", AL = "bmi", COtitle = "Covid", SZtitle = "Age", ALtitle = "BMI", gridTitle = "all but shape")

pcaGrid(screeCumulativeThresholdObject = b, CO = "covid_status", SZ = 5, COtitle = "covid", SZtitle = "Age", gridTitle = "change default size (not to a varaible)")

pcaGrid(screeCumulativeThresholdObject = b, CO = "covid_status", SH = "sex", SZ = "age", AL = 1, COtitle = "covid")

```
#loadings plot grid
```{r}
plotLoadingGrid(screeCumulativeThresholdObject = b)
```
#Multiplot (multiple grids one figure)
#individual PCA for publication 
To contain finer control on colour and shape choice for variable or plain aesthetics
#clustering (kmeans)
